LAI workflow documentation

### 2021-01-29 LAI.Rmd
1. Added new gap-filling approach, prioritising genus and family level over site-level and LMA based rel 2. addeda graph with color-coding by the source of leaf-lifespan

1. Built leaf lifespan (LL) to LMA relationship for San-Lorenzo (FTS) based on canopy leaves and those with sufficient sample size to determine (n_lifespan > 100) (Figure \@ref(fig:ll-lma)). Per Joe, BCI climate is more similar to San Lorenzo, while PNM site is much drier. More importantly, the BCI 50-ha plot and San Lorenzo support old-growth forests. PNM supports much younger, secondary forest (age unknown, perhaps 100 years). So, used the leaf lifespan-LMA relationship for San Lorenzo. This was used to predict leaf lifespan data for species without leaf lifespan data based on their LMA, after gap-filling species LMA data. For the latter, I used LMA_lamina, which was in turn  gap-filled based on a linear relationship with LMALEAF_AVD or LMADISC_AVD.

@@

    1. For each species, calculated a day of the year average of leaf_fall_LAI (m2/m2), weighted by total trap area of sites observed, to account for differential sampling intensity over time. This gives more weight to dates after 2013 when BCI-50ha site observations began. 
    2. LAI, for the changing fraction on the tree is obtained by 
   looping over days t in 1:365 
    LAI[t] <- sum(leaf_fall_LAI[c(t: (t+sp.lifespan)) %% 365])
    where sp.lifespan are the number of days in species leaf lifespan, named 'sp.lifespan'

```{r LAI-ssn-loop, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop}Mean LAI seasonality curves based on Joe's calculation. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.If LL is estimated from LMA it is shown in solid line, if estimated from observations it is in a broken line.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop.jpeg"))
```

@@

# IGNORE THE REST
6. Old : For each species, prepended its leaf_gm_per_m2 data-frame by adding the number of prior dates equivalent to days in the species leaf lifespan, named 'sp.lifespan'  
    Then LAI time series is calculated with the following steps: For t in 1 to c(total no of days in the data frame - leaf.lifespan),   
    1. Assuming zero LAI to begin with, leaf gain on date 1 is leaf fall for the future date shifted by sp.lifespan. That is,    
    If t == 1,  
    LAI[t] = leaf_fall_LAI[sp.lifespan]  
    1. New leaf gain is added to LAI of the previous date, but leaf fall does not begin until the future date equivalent to sp.lifespan. So,  
    For t > 1, if t < sp.lifespan,   
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan]  
    1. Leaf fall begins on date equivalent to sp.lifespan. So,  
    For t > 1, if t >= sp.lifespan,  
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan] - leaf_fall_LAI[t]  
    1. Because dates before the date when leaf-fall begins, that is, the prepended dates from t = 1: sp.lifespan, would underestimate LAI because of lack of a prior, I removed the time series for the prepended dates. The remaining LAI time series for each species is shown in (Figure \@ref(fig:LAI-ts)).

>>>>

2. Gap-filled species-level mean leaf lifespan, with genus-level mean. If genus-level not available, then family-level mean. When neither were available: (1) for deciduous species, used mean lifespan at PNM site, (2) for evergreen species, used a relationship of leaf lifespan (LL) with LMA. This LMA-based relationship was based on data at San-Lorenzo (FTS) for canopy leaves and for species with sufficient sample sizes (n_lifespan > 100) (Figure \@ref(fig:ll-lma)). Per Joe, BCI climate is more similar to San Lorenzo, while PNM site is much drier. More importantly, the BCI 50-ha plot and San Lorenzo support old-growth forests. PNM supports much younger, secondary forest (age unknown, perhaps 100 years). So, used the leaf lifespan-LMA relationship for San Lorenzo. This was used to predict leaf lifespan data for species without leaf lifespan data based on their LMA, after gap-filling species LMA data:
    Gap-filled LMALAM_AVD with the following sequence of priority: 
    LMALAM_AVD relationship with LMADISC_AVD > LMALAM_AVD rel. with LMALEAF_AVD > Genus-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD) > Family-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD)

@@

    1. For each species, calculated a day of the year average of leaf_fall_LAI (m2/m2), weighted by total trap area of sites observed, to account for differential sampling intensity over time. This gives more weight to dates after 2013 when BCI-50ha site observations began. (Figure \@ref(fig:leaf_fall_LAI))) 
    2. LAI, for the changing fraction on the tree is obtained (Figure \@ref(fig:LAI-ssn-loop)) by   
    looping over days t in 1:365   
    LAI[t] <- sum(leaf_fall_LAI[c(t: (t+sp.lifespan)) %% 365])  
    where sp.lifespan are the number of days in species leaf lifespan, named 'sp.lifespan'  
    To normalise LAI, it is the divided by the maximum value.

```{r leaf_fall_LAI, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{leaf_fall_LAI}Seasonality curves for Leaf Fall in LAI units based on Joe's calculation. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/Leaf_fall_LAI.seasonality_mean.jpeg"))
```

```{r LAI-ssn-loop, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop}Mean LAI seasonality curves based on Joe's approach to calculation. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop.jpeg"))
```

```{r LAI-ssn-loop-subcol, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop-subcol}Mean LAI seasonality curves based on Joe's approach to calculation. LL = species-specific leaf lifespan in days. Species are color coded by the method/source of leaf-lifespan substitution.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop_color_substitution.jpeg"))
```

@@

# IGNORE THE REST
6. Old : For each species, prepended its leaf_gm_per_m2 data-frame by adding the number of prior dates equivalent to days in the species leaf lifespan
    Then LAI time series is calculated with the following steps: For t in 1 to c(total no of days in the data frame - leaf.lifespan),   
    1. Assuming zero LAI to begin with, leaf gain on date 1 is leaf fall for the future date shifted by sp.lifespan. That is,    
    If t == 1,  
    LAI[t] = leaf_fall_LAI[sp.lifespan]  
    1. New leaf gain is added to LAI of the previous date, but leaf fall does not begin until the future date equivalent to sp.lifespan. So,  
    For t > 1, if t < sp.lifespan,   
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan]  
    1. Leaf fall begins on date equivalent to sp.lifespan. So,  
    For t > 1, if t >= sp.lifespan,  
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan] - leaf_fall_LAI[t]  
    1. Because dates before the date when leaf-fall begins, that is, the prepended dates from t = 1: sp.lifespan, would underestimate LAI because of lack of a prior, I removed the time series for the prepended dates. The remaining LAI time series for each species is shown in (Figure \@ref(fig:LAI-ts)).
